---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Jamieson Nathan"
date: "07/02/2025"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r dataset-variables}

dim(dennys) 
head(dennys) 

```
Dimensions are 1643 X 6. Each row represents a specific Denny's location and the columns give the identifying information. 

### Exercise 2

```{r other-variables}

dim(laquinta) 
head(laquinta) 

```

909 X 6, each row represents the location of a Laquinta and the columns give the identifying information. 

### Exercise 3

Looking at the websites the data is scraped from, there appear to be numerous overseas locations for both chains (including Asia, Africa, and Europe). 

### Exercise 4

Filtering both datasets by states seems like the best move.Lets have a look at the states dataset to doublecheck it has abbreviations:

```{r some-other-variables}

dim(states) 
head(states) 

```

### Exercise 5

```{r filter-variables}

dennys %>%
  filter(!(state %in% states$abbreviation))

```

There appear to be no Denny's outside of the US (based on this data scrape).


### Exercise 6

```{r new-variable}

dennys <- dennys %>%
  mutate(dn_country = "United States")

```
 
### Exercise 7

```{r laquinta-variables}

laquinta %>%
  filter(!(state %in% states$abbreviation))

```

### Exercise 8

```{r laquinta-country-variables}

laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb ~ "United States",  # Ensure 'state.abb' is defined or replace it with appropriate state codes.
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT" ~ "Colombia",
    state %in% c("AG", "QR", "CH", "NL", "VE", "PU", "SL") ~ "Mexico", 
    state == "FM" ~ "Honduras"
  )) %>%
  filter(country == "United States")

```

# Exercise 9

```{r frequency}

dennys_count <- dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))

laquinta_count <- laquinta %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))

```

```{r more-frequency}

dennys_most <- filter(dennys_count, n == max(n))
dennys_fewest <- filter(dennys_count, n == min(n))

laquinta_most <- filter(laquinta_count, n == max(n))
laquinta_fewest <- filter(laquinta_count, n == min(n))

print(dennys_most)
print(dennys_fewest)
print(laquinta_most)
print(laquinta_fewest)

```

Cali/Delaware have the most/fewest Denny's locations respectively, whereas, Texas/Maine have the same for Laquinta. Not too surprising, given Ca And Tx are typically the most popular states and the other two are much less populated. 

# Exercise 10

```{r density}

dennys_density <- dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(density_per_thousand_sq_mi = n / (area / 1000))  

dennys_most_dense <- dennys_density %>%
  arrange(desc(density_per_thousand_sq_mi)) %>%
  slice(1)

dennys_least_dense <- dennys_density %>%
  arrange(density_per_thousand_sq_mi) %>%
  slice(1)

laquinta_density <- laquinta %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(density_per_thousand_sq_mi = n / (area / 1000))  

laquinta_most_dense <- laquinta_density %>%
  arrange(desc(density_per_thousand_sq_mi)) %>%
  slice(1)

laquinta_least_dense <- laquinta_density %>%
  arrange(density_per_thousand_sq_mi) %>%
  slice(1)

print(dennys_most_dense)
print(dennys_least_dense)
print(laquinta_most_dense)
print(laquinta_least_dense)

```
This makes sense, as DC and RI are incredible dense states, so it tracks that it is the highest and Alaska is the least for both

```{r combining}

dn <- dennys %>%
  mutate(establishment = "Denny's")
lq <- laquinta %>%
  mutate(establishment = "La Quinta")

dn_lq <- bind_rows(dn, lq)


```

```{r plot}

usa_locations_plot <- ggplot(dn_lq, aes(
  x = longitude,
  y = latitude,
  color = establishment
)) +
  geom_point(alpha = 0.5, size = 2) +  
  labs(title = "Denny's and La Quinta Locations Across the USA",
       x = "Longitude",
       y = "Latitude",
       color = "Establishment Type") +
  theme_minimal() +
  theme(legend.position = "bottom")  
print(usa_locations_plot)

```
# Exercise 11

```{r NC}

nc_locations <- dn_lq %>%
  filter(state == "NC")

nc_locations_plot <- ggplot(nc_locations, aes(
  x = longitude,
  y = latitude,
  color = establishment
)) +
  geom_point(alpha = 0.4, size = 3) +  
  labs(title = "Denny's and La Quinta Locations in North Carolina",
       x = "Longitude",
       y = "Latitude",
       color = "Establishment Type") +
  theme_minimal() +
  theme(legend.position = "bottom")  

print(nc_locations_plot)

```

Partially, but there are enough stand alones to see some variability in the joke. Although it does seem that where there is a La Quinta, there is a Dennys' for the most part. 


# Exercise 11

```{r TX}

tx_locations <- dn_lq %>%
  filter(state == "TX")

tx_locations_plot <- ggplot(tx_locations, aes(
  x = longitude,
  y = latitude,
  color = establishment
)) +
  geom_point(alpha = 0.3, size = 2) +  
  labs(title = "Denny's and La Quinta Locations in Texas",
       x = "Longitude",
       y = "Latitude",
       color = "Establishment Type") +
  theme_minimal() +
  theme(legend.position = "bottom")  

print(tx_locations_plot)

```

Now he seems to be pretty spot on! It is almost uncanny how close these two chains follow each other here. He must be bilingual. 