---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Jamieson Nathan"
date: "07/02/2025"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r dataset-variables}

dim(dennys) 
head(dennys) 

```
Dimensions are 1643 X 6. Each row represents a specific Denny's location and the columns give the identifying information. 

### Exercise 2

```{r other-variables}

dim(laquinta) 
head(laquinta) 

```

909 X 6, each row represents the location of a Laquinta and the columns give the identifying information. 

### Exercise 3

Looking at the websites the data is scraped from, there appear to be numerous overseas locations for both chains (including Asia, Africa, and Europe). 

### Exercise 4

Filtering both datasets by states seems like the best move.Lets have a look at the states dataset to doublecheck it has abbreviations:

```{r some-other-variables}

dim(states) 
head(states) 

```

### Exercise 5

```{r filter-variables}

dennys %>%
  filter(!(state %in% states$abbreviation))

```

There appear to be no Denny's outside of the US (based on this data scrape).


### Exercise 6

```{r new-variable}

dennys %>%
  mutate(dn_country = "United States")

```
 
### Exercise 7

```{r laquinta-variables}

laquinta %>%
  filter(!(state %in% states$abbreviation))

```

### Exercise 8

```{r laquinta-country-variables}

laquinta %>%
  mutate(country = case_when(
    state %in% state.abb ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT" ~ "Colombia",
    state %in% c("AG", "QR", "CH", "NL", "VE", "PU", "SL") ~ "Mexico", 
    state == "FM" ~ "Honduras"
  ))

laquinta <- laquinta %>%
  filter(country == "United States")

```

```{r filter-for-USA}


```


